\chapter{Operational Scenarios \\
\small{\textit{-- Emmanuel Okoro and Spurthi Setty }}
\index{Operational Scenarios} 
\index{Chapter!Operational Scenarios}
\label{Chapter::Operational Scenarios}}

\section{Operational Scenarios}

\subsection*{Scenario 1: First-time User Setup}
\begin{itemize}
    \item \textbf{Stimulus:} A new user installs the app and sets up an account.
    \item \textbf{Response:} The app prompts the user to link a bank account for automated financial tracking or manually enter initial financial data. 
    \item \textbf{Outcome:} A personalized dashboard is created, automatically importing recent bank transactions to provide an initial financial snapshot. Users are also guided to set up receipt scanning and tax assistance preferences.
\end{itemize}

\subsection*{Scenario 2: Monthly Budget Review}
\begin{itemize}
    \item \textbf{Stimulus:} User views the monthly spending summary and budget performance.
    \item \textbf{Response:} The app displays detailed analytics, highlights categories where spending exceeded the budget, and provides real-time recommendations to optimize financial goals.
    \item \textbf{Outcome:} User adjusts spending habits, sets new savings goals, and receives actionable insights to maintain better financial health.
\end{itemize}

\subsection*{Scenario 3: Tax Filing Season}
\begin{itemize}
    \item \textbf{Stimulus:} User initiates the tax assistance tool during tax season.
    \item \textbf{Response:} The app scans user-provided financial data and receipts, calculates tax obligations, and pre-fills tax forms based on user input and uploaded documents.
    \item \textbf{Outcome:} User reviews and submits tax returns with minimal effort, benefiting from accurate calculations and compliance with tax regulations.
\end{itemize}

\subsection*{Scenario 4: Receipt Scanning for Expense Tracking}
\begin{itemize}
    \item \textbf{Stimulus:} User uploads a receipt after making a purchase.
    \item \textbf{Response:} The app uses AI to extract item details, prices, and total amounts from the receipt, categorizing the expenses automatically.
    \item \textbf{Outcome:} The transaction is added to the user's financial records without manual data entry, improving accuracy and efficiency.
\end{itemize}

\subsection*{Scenario 5: Investment Tracking and Insights}
\begin{itemize}
    \item \textbf{Stimulus:} User accesses the investment tracking feature to review portfolio performance.
    \item \textbf{Response:} The app provides detailed investment analytics, including growth trends, risk levels, and forecasting tools, along with real-time recommendations for portfolio optimization.
    \item \textbf{Outcome:} User gains actionable insights to make informed investment decisions and improve long-term financial outcomes.
\end{itemize}