\chapter{Summary of Impacts \\
\small{\textit{-- Emmanuel Okoro and Spurthi Setty}}
\index{Summary of impacts} 
\index{Chapter!\textbf{Summary of Impacts}}
\label{Chapter::Summary of Impacts}}

\section{Operational Impacts}

\subsection{Impacts on Users}
\begin{itemize}
    \item \textbf{Interfaces with Primary or Alternate Systems:} Users will interface with the app through mobile and web platforms, enabling seamless access to financial data in real-time. Emergency backup systems ensure data availability during outages.
    \item \textbf{Changes in Procedures:} Users will transition from manual financial tracking to automated data input and analysis, requiring minimal adjustment to existing workflows.
    \item \textbf{Use of New Data Sources:} The app introduces integration with financial institutions and receipt scanning for automated data collection, enhancing the accuracy and scope of input data.
    \item \textbf{Changes in Data Quantity and Timing:} Users can expect more frequent and comprehensive data updates, offering better insights into financial trends.
    \item \textbf{Changes in Data Retention Requirements:} Data will be securely stored and retained in compliance with regulatory requirements, with user-defined retention policies for specific financial data.
    \item \textbf{New Operational Modes:} Emergency data recovery mechanisms ensure operational continuity in disaster scenarios.
    \item \textbf{New Input Methods:} Receipt scanning and AI-driven data extraction provide alternative methods for users to input financial data when direct integrations are unavailable.
    \item \textbf{Changes in Operational Budget:} The app helps users optimize spending and budgeting, indirectly reducing financial stress and improving cash flow management.
    \item \textbf{Operational Risks:} Potential risks include data breaches or inaccuracies in AI-generated insights, mitigated by robust encryption and ongoing model refinement.
\end{itemize}

\subsection{Impacts on Support or Maintenance Agencies}
\begin{itemize}
    \item \textbf{Changes in Responsibilities:} Support teams will focus on maintaining app functionality, addressing user queries, and ensuring seamless updates.
    \item \textbf{Training Requirements:} Maintenance staff may require additional training in cloud infrastructure management, AI monitoring, and data privacy regulations.
    \item \textbf{Personnel Allocation:} Increased reliance on cloud services and data analytics might necessitate hiring or upskilling personnel in relevant domains.
\end{itemize}

\section{Organizational Impacts}

\subsection{Impacts on Development and Operations}
\begin{itemize}
    \item \textbf{Modification of Responsibilities:} Financial institutions may need to adjust their data-sharing protocols to integrate with the app.
    \item \textbf{Job Positions:} Organizations might create or eliminate roles related to manual financial tracking due to automation.
    \item \textbf{Training or Retraining:} Staff will require training on how to effectively use and support the app’s advanced features.
    \item \textbf{Changes in Personnel Requirements:} Skilled personnel in AI, cloud computing, and cybersecurity will be required to support app operations.
    \item \textbf{Contingency Operations:} Emergency procedures will need to be established to ensure data recovery and system continuity in case of disasters.
\end{itemize}

\section{Impacts During Development}

\subsection{Involvement of Users and Stakeholders}
\begin{itemize}
    \item \textbf{Pre-Award Activities:} Users and stakeholders will participate in requirements elicitation, feasibility studies, and system specification reviews.
    \item \textbf{Reviews and Demonstrations:} Users will provide feedback during prototype reviews and system demonstrations to ensure the app meets their expectations.
    \item \textbf{Database Development:} Development teams will work with users to populate and validate financial data databases during the development phase.
\end{itemize}

\subsection{Operational Testing and Transition}
\begin{itemize}
    \item \textbf{Parallel Operation:} During system testing, the app may operate in parallel with existing systems to ensure a smooth transition and identify any discrepancies.
    \item \textbf{System Testing Impacts:} Users may need to participate in testing scenarios, such as verifying data accuracy and performance under different conditions.
\end{itemize}


